Formalization of Forcing in Isabelle/ZF

### 2021-05-28 IJCAR2020/slides/forcing_IJCAR2020_slides.tex

  Every $a\in M[G]$ is coded by
  some $\punto{a}\in M$ through the function $\val$:
  \[
  M[G]\defi \{ \val(G,\punto{a}) : \punto{a}\in M\}
  \]
  Fundamentally, \textbf{truth} in $M[G]$ is coded in $M$ by the
  function $\forceisa$.
  \pause
  \begin{theorem}[\citep{Cohen-CH-PNAS}]
    \begin{enumerate}
    \item $\forceisa$ is definable over $M$.
      %
    \item For every $\phi$, $M$-generic $G$, and $\punto{a}\in M$,
      \[
      M[G], [\val(G,\punto{a})] \models \phi
      \quad      \iff  \quad 
      \exists p\in G.\ \; M, [p,\preceq,\PP, \punto{a}]  \models
      \forceisa(\phi).
      \]
      \pause
      \rule{21em}{0pt}
      \raisebox{1.6ex}[0ex]{$\overset{\searrow}{}$ \ \ 
        $p \alert{\forces^M_{\PP,\preceq}} \phi(\punto{a})$
        \  \ $\overset{\swarrow}{}$}
    \end{enumerate}
  \end{theorem} 
\end{frame}

\begin{frame}
  \frametitle{The ctm approach to forcing, 3/4}
  \begin{block}{Benefits of using ctms}
    \begin{enumerate}
    \item<+-> Countability ensures that generics exist (by
      Rasiowa-Sikorski Lemma).
      %  
    \item<+-> Absoluteness provides seamless
      treatment of many concepts.
      \begin{center}
        $\alpha$ is an ordinal \ $\iff$ \  $M\models \alpha$ is
        an ordinal \  $\iff$ \  $M[G]\models \alpha$ is an ordinal
      \end{center}
      %
    \item<+-> Both $M$ and $M[G]$ are standard (two-valued) models.
      %
    \item<+-> Ctms are
      used in an important (perhaps the greatest) fraction of the literature.
    \end{enumerate}
  \end{block}
\end{frame}

@@

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%
\begin{frame}
  \begin{shadowblock}{}
    \begin{center}
      {\Huge Thank you!}
    \end{center}
  \end{shadowblock}
\end{frame}

\begin{frame}
  \frametitle{References}
  \bibliographystyle{mi-estilo-else}
  \bibliography{../forcing_in_isabelle_zf}
\end{frame}

>>>>

  Every $a\in M[G]$ is coded by
  some $\punto{a}\in M$ through the function $\val$:
  \[
  M[G]\defi \{ \val(G,\punto{a}) : \punto{a}\in M\}
  \]
  Fundamentally, \textbf{truth} in $M[G]$ is coded in $M$ by the
  function $\forceisa$.
  \pause
  \begin{theorem}[\citep{Cohen-CH-PNAS}]
    There exists a formula-transformer  $\forceisa$
    such that for every $\phi$, $M$-generic $G$, and $\punto{a}\in M$,
      \[
      M[G], [\val(G,\punto{a})] \models \phi
      \quad      \iff  \quad 
      \exists p\in G.\ \; M, [p,\preceq,\PP, \punto{a}]  \models
      \forceisa(\phi).
      \]
      \pause
      \rule{21em}{0pt}
      \raisebox{1.6ex}[0ex]{$\overset{\searrow}{}$ \ \ 
        $p \alert{\forces^M_{\PP,\preceq}} \phi(\punto{a})$
        \  \ $\overset{\swarrow}{}$}
  \end{theorem} 
\end{frame}

\begin{frame}
  \frametitle{The ctm approach to forcing, 3/4}
  \begin{block}{Benefits of using ctms}
    \begin{enumerate}
    \item<+-> Countability ensures that generics exist (by
      Rasiowa-Sikorski Lemma).
      %  
    \item<+-> Absoluteness provides seamless
      treatment of many concepts.
      \begin{center}
        $\alpha$ is an ordinal \ $\iff$ \  $M\models \alpha$ is
        an ordinal \  $\iff$ \  $M[G]\models \alpha$ is an ordinal
      \end{center}
      %
    \item<+-> Both $M$ and $M[G]$ are standard (two-valued) models.
      %
    \item<+-> Ctms are
      used in an important fraction of the literature.
    \end{enumerate}
  \end{block}
\end{frame}

@@

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%
\begin{frame}
  \begin{shadowblock}{}
    \begin{center}
      {\Huge Thank you!}
    \end{center}
  \end{shadowblock}
  \transduration{2}
\end{frame}

\begin{frame}
  \frametitle{References}
  \bibliographystyle{mi-estilo-else}
  \bibliography{../forcing_in_isabelle_zf}
  \transduration{5}
\end{frame}
